Morse Theory

### 2020-12-06 paper.tex
final update

% \doublespacing
\begin{document} 
\maketitle

\begin{figure}[h]
    \centering
    \includegraphics[width=0.50\textwidth]{cover}
\end{figure}

\tableofcontents
\newpage

\begin{abstract}
    In this paper, we studied Morse functions and introduction to Morse theory. Morse Theory allows us to conduct calculus not only on simple manifolds (i.e. planes), but also on complicated ones as Klein Bottle. Therefore, it is obligatory to research the theory. Here we will prove Morse Lemma, give examples of Morse functions. Ideas of Morse Theory are also used for Reeb Graphs. In this work, we will introduce Reeb graphs, define their properties on orientable and non-orienable 2-manifolds. Reeb graphs are widely used in many applications as topological data analysis, computer graphics, oceanography. Main ideas of these applications will also be discussed. 
\end{abstract}

A smooth function is always good to have. However, some critical points might have equal function values. Morse functions are such ones that don't have the latter property. 
Sometimes we would like to analyze a function, using its level sets. A Reeb graph helps us plot the evolution of the components of a level set. Morse functions and Reeb graphs are extremely useful
in some contexts. They have applications in topological data analysis, computer graphics, and even in oceanography.
\section{Morse Theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % begin N1
\subsection{Introduction}

\begin{definition}(Morse function)
A smooth function on a manifold is a Morse function if 
\begin{itemize}
	\item all critical points are non-degenerate
	\item the critical points have distinct function values
\end{itemize}
\end{definition}

@@

\subsection{Examples of Morse functions} 
\begin{problem}
Show that the function $f: S^2 \to \R, f(x,y,z) = z$ is a Morse function.
\end{problem}
$f$ is smooth on $ S^2 $ since it extends to a smooth map on all of $\R^3$. We can map $R^3$ to $S^2$ with the stereographic projection two functions:
\begin{equation}
	\phi_1(x_1,x_2,x_3) = (\frac{x_1}{1-x_3}, \frac{x_2}{1-x_3} ) 
\end{equation}
and
\begin{equation}
	\phi_2(y_1,y_2,y_3) = (\frac{y_1}{1+y_3},\frac{y_2}{1+y_3})
\end{equation}
Therefore, to get $S^2 \to R^3$ we can take $\phi_1^{-1}$ and $\phi_2^{-1}$ \\
\begin{equation}
\phi_1^{-1}(x_1,x_2,x_3) = (\frac{2x_1}{x_1^2+x_2^2+1}, \frac{2x_2}{x_1^2+x_2^2+1}, \frac{x_1^2+x_2^2-1}{x_1^2+x_2^2+1})
\end{equation}
\begin{equation}
	\phi_2^{-1}(y_1,y_2,y_3) = (\frac{2y_1}{y_1^2+y_2^2+1}, \frac{2y_2}{y_1^2+y_2^2+1}, \frac{1-y_1^2+y_2^2}{y_1^2+y_2^2+1})
\end{equation}
Then, we take $g_1 = f \circ \phi_1^{-1}$ and $g_2 = f \circ \phi_2^{-1}$ \\
\begin{equation}
	g_1(x,y) =  \frac{x^2+y^2-1}{x^2+y^2+1}
\end{equation}
and
\begin{equation}
	g_2(x,y) = \frac{1-x^2+y^2}{x^2+y^2+1}
\end{equation}
Now, we can compute jacobian of $g_1$ and $g_2$, find critical points, and check that determinant of hessian matrix is non-zero. \\
$
	\nabla g_1(x,y,z) = 0$ iff $(x,y,z) = (0,0,-1)
$\\
$
	\nabla g_2(x,y,z) = 0$ iff $ (x,y,z) = (0,0,1)
$
We have two critical points, and now we compute the hessian at them.\\

>>>>

% \doublespacing
\begin{document} 
\maketitle
\begin{abstract}
    In this paper, we studied Morse functions and introduction to Morse theory. Morse Theory allows us to conduct calculus not only on simple manifolds (i.e. planes), but also on complicated ones as Klein Bottle. Therefore, it is obligatory to research the theory. Here we will prove Morse Lemma, give examples of Morse functions. Ideas of Morse Theory are also used for Reeb Graphs. In this work, we will introduce Reeb graphs, define their properties on orientable and non-orienable 2-manifolds. Reeb graphs are widely used in many applications as topological data analysis, computer graphics, oceanography. Main ideas of these applications will also be discussed. 
\end{abstract}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.50\textwidth]{cover}
\end{figure}

\newpage
\tableofcontents
\newpage

\section{Morse Theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % begin N1
\subsection{Introduction}

A smooth function is always good to have. However, some critical points might have equal function values. Morse functions are such ones that don't have the latter property. 
Sometimes we would like to analyze a function, using its level sets. A Reeb graph helps us plot the evolution of the components of a level set. Morse functions and Reeb graphs are extremely useful
in some contexts. They have applications in topological data analysis, computer graphics, and even in oceanography.
\

\begin{definition}(Morse function)

A smooth function on a manifold is a Morse function if 
\begin{itemize}
	\item all critical points are non-degenerate
	\item the critical points have distinct function values
\end{itemize}
\end{definition}

@@

\subsection{Examples of Morse functions} 
\begin{example}
Show that the function $f: S^2 \to \R, f(x,y,z) = z$ is a Morse function.
\end{example}
$f$ is smooth on $ S^2 $ since it extends to a smooth map on all of $\R^3$. We can map $R^3$ to $S^2$ with the stereographic projection two functions:
\begin{equation}
	\phi_1(x_1,x_2,x_3) = (\frac{x_1}{1-x_3}, \frac{x_2}{1-x_3} ) 
\end{equation}
and
\begin{equation}
	\phi_2(y_1,y_2,y_3) = (\frac{y_1}{1+y_3},\frac{y_2}{1+y_3})
\end{equation}
Therefore, to get $S^2 \to R^3$ we can take $\phi_1^{-1}$ and $\phi_2^{-1}$ \\
\begin{equation}
\phi_1^{-1}(x_1,x_2,x_3) = (\frac{2x_1}{x_1^2+x_2^2+1}, \frac{2x_2}{x_1^2+x_2^2+1}, \frac{x_1^2+x_2^2-1}{x_1^2+x_2^2+1})
\end{equation}
\begin{equation}
	\phi_2^{-1}(y_1,y_2,y_3) = (\frac{2y_1}{y_1^2+y_2^2+1}, \frac{2y_2}{y_1^2+y_2^2+1}, \frac{1-y_1^2+y_2^2}{y_1^2+y_2^2+1})
\end{equation}
Then, we take $g_1 = f \circ \phi_1^{-1}$ and $g_2 = f \circ \phi_2^{-1}$ \\
\begin{equation}
	g_1(x,y) =  \frac{x^2+y^2-1}{x^2+y^2+1}
\end{equation}
and
\begin{equation}
	g_2(x,y) = \frac{1-x^2+y^2}{x^2+y^2+1}
\end{equation}
Now, we can compute jacobian of $g_1$ and $g_2$, find critical points, and check that determinant of hessian matrix is non-zero. \\
$
	\nabla g_1(x,y,z) = 0$ iff $(x,y,z) = (0,0,-1)
$\\
$
	\nabla g_2(x,y,z) = 0$ iff $ (x,y,z) = (0,0,1)
$
We have two critical points, and now we compute the hessian at them.\\
